---
title: The Impact of NBA player-related Social Media Posts on their on-court Performance
  - An Analysis
author: "Frank Dreyer, Kolja Günther, Jannik Greif"
date: "20.05.2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: DataSciR - Project Notebook
bibliography: references.bib
csl: ieee.csl
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center} \includegraphics[width=2in,height=2in]{"datascir.png"}\LARGE\\}
- \posttitle{\end{center}}
link-citations: yes
---
\newpage

## Github Repository

The project is documented here: https://github.com/jannikgreif/DataSciR_2021


## Team Members

```{r team, echo = FALSE}
library(knitr)
library(kableExtra)
team <- data.frame(Name = c("Jannik Greif","Kolja Günther","Frank Dreyer"),
                   `Course of Studies` =c("M.Sc. Wirtschaftsinformatik","M.Sc. Data and Knowledge Engineering","M.Sc. Data and Knowledge Engineering"),
                  Mail = c("jannik.greif@st.ovgu.de","kolja.guenther@st.ovgu.de","frank.dreyer@st.ovgu.de"), check.names = FALSE)
kbl(team, booktabs = T, linesep = "") %>%
  kable_styling(position = "center")%>%
  kable_styling(latex_options = "HOLD_position")
```

## Overview
The project aims to discover a significant impact of social media posts addressed to NBA players before matches with respect to their influence on these players’ game performance. For this purpose, we consider NBA players that are highly active on Twitter and extract tweets that are addressed to them within a short period of time before matches via the Twitter API. A sentiment analysis indicates the attitude of the posts and with the resulting sentiment polarity scores we test if there is a correlation between social media posts and players’ on-court performance.

## Background and Motivation
With the growing presence of social media in all areas of life, allowing people from around the world to react to current events in real-time, an increasingly controversial discussion can be noticed. Today more than ever, public figures are exposed to the reactions of millions of people, observing and commenting on every step in their life that becomes public. The resulting negative impact that extensive social media usage can have on users' behavior and mental state is subject to different scientific studies [-@kapoor_advances_2018;-@berryman_social_2018]. 
 
Sports athletes, who use social media not only to communicate with peers and fans but also to promote themselves, are no exception to these issues [-@academy_does_2008]. Among researchers in the sports field, there is a consensus that the mental state of an athlete can have a significant impact on his or her performance [-@xu_measuring_2015]. However, only little research has been conducted in order to analyze if and how social media usage of athletes directly influences their performance.
Xu and Yu [-@xu_measuring_2015] tried to capture the mood of basketball players in the NBA from the tweets they posted just before a match, using sentiment analysis, to analyze how the predicted mood influenced their performance on court. Gruettner, Vitisvorakarn and Wambsganss [-@gruttner_new_2020] used a similar approach on tennis players and additionally analyzed the relationship between the number of tweets they posted before matches and their performance within the match. Even though both contributions show that athletes with a bad predicted mood tend to perform worse on-court, they suffer from two limitations: 
 
- The number of tweets an athlete posts per day is rather limited 
- The predicted moods are not free of bias since an athlete might only post tweets how he or she wants to be seen on Twitter (also indicated in [-@gruttner_new_2020])
 
Both of these limiting factors may lead to an inaccurate prediction of the mental state of athletes. \newline
We believe that the attitude of social media posts an athlete receives from peers and fans is also a good predictor for his or her performance. Ott and Puymbroeck support this claim  [-@academy_does_2008]. In their article they list cases where athletic performance appeared to be immediately influenced by the media and conclude that the media has the potential to change the performance of an athlete in a negative as well as positive way. In this analysis we aim to assess this relationship more closely by analyzing how social media posts addressed to NBA players affect their game performance. 

## Related Work

\newpage

## Initial Questions and Project Objectives

This project aims to answer the following research question: \newline
**Does the attitude of social media posts addressed to NBA players affect their performance in games?** \newline

To answer the specified research question the following objectives can be formulated: \newline

**Objective 1: Dataset Creation** \newline
Acquire game statistics of NBA players that are highly active on Twitter and the tweets they received from peers and fans in an appropriate time window before games. The game statistics should include an appropriate metric that describes how the player performed within a corresponding game. The tweets need to be preprocessed accordingly to have them in an appropriate format in order to use them for further analysis steps. The attitude of the extracted posts should be captured by assigning a sentiment score to them. The sentiment scores of the tweets a player received in the corresponding time window before a game should be aggregated accordingly and linked to the respective game. As a result, this should end in a data set in which each record contains the game statistics of a player for a specific game and the aggregated sentiment information of the tweets that were addressed to the player before the game. 

**Objective 2: Exploratory Data Analysis** \newline
Analyze the association between the aggregated polarity scores of the tweets a player received before games and the performance of the player within the games using appropriate performance metrics. Additionally, the strength and significance of the correlation should be evaluated. 

**Objective 3: Presentation ** \newline
Document the implementation and the results of the analysis in a document. Additionally, the results are summarized in a screencast and a web application.
 
With this, we can refine our initial research question as follows: \newline
**Can we find a correlation between negative/positive Social Media posts related to a specific NBA player and his on-court performance in the following matches?**

## The Datasets

**Twitter API/ rtweet:** \newline    
In order to access the Twitter API, it is a prerequisite to create a Twitter account and apply for a developer account. After the acceptance, it is required to register an app to generate the API keys [-@noauthor_twitter_nodate-1]. To access the Twitter data in R, we are using the package rtweet [-@kearney_collecting_2020]. 
Important functionalities are:
```{r twitter api, echo = FALSE}
library(knitr)
funtions <- data.frame(Functionality = c("create_token","get_mentions", "get_timeline", "lookup_users", "parse_stream","plain_tweets","search_tweets", "stopwordslangs", "write_as_csv"), 
                   Explanation = c("Creating Twitter authorization token(s)",
                                   "Get mentions for the authenticating user",
                                   "Get one or more user timelines (tweets posted by target user(s))",
                                   "Get Twitter users data for given users (user IDs or screen names)",
                                   "Converts Twitter stream data (JSON file) into parsed data frame",
                                   "Clean up character vector (tweets) to more of a plain text",
                                   "Get tweets data on statuses identified via search query",
                                   "Twitter stop words in multiple languages data",
                                   "Save Twitter data as a comma separated value file") )
kbl(funtions, booktabs = T, linesep = "") %>%
  kable_styling(position = "center") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_styling(latex_options = "striped")
```
\newpage

\newpage

**NBA players on-court performance dataset and Twitter accounts:** \newline
[basketball-reference.com](basketball-reference.com) provides historical basketball statistics about different players and teams in various US-American and European leagues. 
From this site, we aim to extract game statistics from NBA players that are active on Twitter. 
Additionally, basketball-reference offers a list of [Twitter accounts](https://www.basketball-reference.com/friv/twitter.html)  from various players that we want to incorporate in this process.
The statistics of a specific game of a player can be reached by clicking on the players’ name (e.g. from the Twitter account list), selecting a season the player was active (regular season or playoffs), and finally selecting a game within this season by clicking on its date. Then a table shows up that shows the player statistics of the selected game. 
This table provides the following information: 

```{r basketball-reference, echo = FALSE}
library(knitr)
table <- data.frame(
  Attribute = c("Starters / Reserves","MP","FG","FGA","FG%","3P","3PA","3P%","FT","FTA","FT%","ORB","DRB","TRB","AST","STL","BLK","TOV","PF","PTS","+/-"),
  `Data Type` = c("String","Timediff","Int","Int","Float","Int","Int","Float","Int","Int","Float",rep("Int",times=10)),
  Description = c("Name of player (separated in starters and reserves)","Minutes Played","Field Goals: number made shots (excluding free throws)","Field Goal Attempts = number of shot attempts (excluding free throws)","Field Goal Percentage: fraction of field goal attempts (FG/FGA)","3-Point Field Goals: number of made 3-point shots","3-Point Field Goal Attempts: number of 3-point shot attempts","3-Point Field Goal Percentage: fraction of three point shot attempts (3P/3PA)","Free Throws: number of free throw shots ","Free Throw Attempts: number of free throw shot attempts","Free Throw Percentage: fraction of free throw attempts (FT/FTA)","Offensive Rebounds","Defensive Rebounds","Total Rebounds (ORB+TRB)","Assists","Steals","Blocks","Turnovers","Personal Fouls","Points made","Estimates the players’ contribution to the team when the player is on the court"), check.names = FALSE
)
kbl(table, booktabs = T, linesep = "") %>%
  kable_styling(position = "center") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_styling(latex_options = "striped")
```

## The Data

<!--To answer the specified research question we used the statistical programming language R including the following packages: 

- tidyverse (includes packages like ggplot2 for visualization, dplyr for data manipulation, etc.) [-@tidyverse]
- rvest [-@rvest]
- rtweet [@kearney_collecting_2020]
- tidytext [-@tidytext]
- Shiny [-@shiny]

We addressed the project objectives as follows: \newline

**Objective 1: Dataset Creation**
--> 
The main challenge in the preprocessing phase of our project was to create suitable data sets which reflect both of our variables we wanted to include into our analysis. For the player's performance variable we needed to create a set of datasets which cover all neccessary statistics and metadata to be able to derive all needed values. For the sentiment variable of the tweets referring to one respective player, we first needed to narrow down the selection of players whose tweets we wanted to contain and then extract all tweets that are related to this set of players. How this was done is described in the following section.

### NBA Stats Datasets
To get the needed data about players, games, seasons and all relevant meta-data, we extracted statistical data from [basketball-reference.com] (https://www.basketball-reference.com),a site which provides historized basketball statistics of basketball players and teams from various US American and European leagues including the NBA. From this we created local .csv files for different metrics.

#### player-metadata
Before extracting stats about NBA players and games, we had to check, which players even have a twitter account. Fortunately for us, [basketball-reference.com](https://basketball-reference.com) provides a list of [Twitter usernames of NBA players](https://www.basketball-reference.com/friv/twitter.html), so we loaded the account names into the player-metadata.csv, along with an unique BBRef_Player_ID, which we took over from [basketball-reference.com](https://basketball-reference.com), and the clear name of the respective players.
With this set of players we now wanted to extract further statistics.

#### player season-stats
The goal of this data set was to create a table which included all the statistics of players on season-level. As a basketball season is split into a regular season (comparable to our "Bundesliga"-system) and a playoff season (comparable to a tournaments k.o.-phase) which only the best teams of one regular season pass to, [basketball-reference.com](https://basketball-reference.com) provides two separate data sets, one for each season type.
To combine those data sets and map them to each player in one table, we 
Finally the data set contained one tuple of statistics for each player and season/seasontype he participated in, including the following metrics:
- Age: The player's age at the time of the respective season
- Team: The team for which the player was under contract in the respective season
- Pos: The player's position in the team
- G: The number of games, played by the player during the respective season
- GS: The number of games, in which the player started from the beginning on
- MP: The average number of minutes played per game
- FG: The average number of field goals scored per game
- FGA: The average number of field goal attempts per game
- FG%: THe percentage of scored field goals
- 3P: The average number of 3-point-field goals scored per game
- 3PA: The average number of 3-point-field goal attempts per game
- 3P%: The percentage of scored 3-point-field goals
- 2P: The average number of 2-point-field goals scored per game
- 2PA: The average number of 2-point-field goal attempts per game
- 2P%: The percentage of scored 2-point-field goals 
- eFG5: effective field goal percentage, which adjusts for the fact that a 3-point field goal is worth more than a 2-point field goal etc.
- FT: The average number of free Throws per game
- FTA: The average number of free throw attempts per game
- FT%: The average number of free throw percentage
- ORB: The average number of offensive rebounds per game
- DRB: The average number of defensive rebounds per game
- TRB: The average number of total rebounds per game
- AST: The average number of assist per game
- STL: The average number of steals per game
- BLK: The average number of blocks per game
- TOV: The average number of turnovers per game
- PF: The average number of personal fouls per game
- PTS: The average number of points scored per game

#### game-metadata

#### player-game-stats

### Twitter Dataset
The Twitter-Dataset contains all Tweets related to the players we want to inspect in our analysis for their on-court performance and this part contains the whole pipeline of extracting these relevant tweets. But before we can exploit the API, some pre-work has to be done.
First, we set up the main directory for our data to be created and the four datasets we previously created from the NBAStats source get loaded. In the next step and before we could start with the extraction of relevant tweets for our NBA players, we had to narrow down the number of players to be considered by our pipeline. The data sets above were created over 227 players. As the process of collecting tweets for such a number of individuals would be a huge overload, we decided to pick those top players, which are most relevant for us, following some criteria we set up in the following.

#### Players who played actively for the same team
First of all we picked those players, which continuously played in the regular seasons 2016/17 - 2018/19. We didn't consider the playoffs here, as many players didn't get into the playoffs with their teams but still played a full regular season and therefore are provide enough interesting play-data for our analysis. Furthermore, we only wanted those players in our data set, which stayed at their respective team for the whole time of observation. The idea behind this was to eliminate team switches as possible factors that influence the player's performance aside those we want to observe. Additionally we wanted only those players who had on-court time in at least 80% of the games during their regular season.

#### Players whose BPM varied by a standard deviation of at least 8
As third parameter we inspected the variable "Box Plus/Minus" (BPM) in the player_game_stats data set, which is a score-based performance indicator that "estimates a basketball player's contribution to the team when that player is on the court. [...] BPM uses a player's box score information, position, and the team's overall performance to estimate the player's contribution in points above league average per 100 possessions played."[insert reference: https://www.basketball-reference.com/about/bpm2.html]
With this estimate, we wanted to extract those players, whose performance is relatively unstable in comparison to their colleagues by computing the standard deviation of performance for each player and storing them from the highest deviation in descending order. on this data set we applied a cutoff value to get only those players, whose standard deviation was higher or equal to 8.

#### Players with at least 1000 followers
The last parameter we wanted to include into our selection concerned about the players who have a minimal follower count of 1000 users on Twitter. For this we ...!!!

#### Merging the parameters
Finally we merged the cut off standard deviation values of the players with their respective twitter-account data, including the count of followers, the count of posted statuses, the count of accounts indicated as favourites and the player's screen name.Now the last step was to create a final set of players we wanted to consider in our analysis by merging the two data sets created into one and picking the top intersecting players.

### Extracting the relevant tweets
With the given data we were now able to extract exactly those tweets we needed for our analysis. To do so, we chose to use the get_all_tweets function from the academictwitteR packagage.

To extract the tweets we used rtweet which provides access to the Twitter API.
To extract the game statistics of NBA players from basketball-reference.com we will make use of the web scraping package rvest.
Since the measured performance of a player is rather unreliable, if he only gets a small amount of playing time we will only consider players that on average get at least 20 minutes of playing time. 
The performance of players is rather unstable over many seasons. To address this issue we will only consider games in the span of two seasons for a player. 
To make sure that the players selected for our analysis actually read tweets they receive from peers and fans we will analyze the number of daily received tweets to which they react (e.g. received tweets they like/dislike, answer, etc.) and select a subset of the top twitter-active NBA players according to this measure. 
Since the NBA prohibits social media usage of players and coaches beginning 45 minutes before tip-off [-@noauthor_nba_2009], we will consider all tweets that a player receives before that time in a 12 hour window to be able to analyze the immediate effects of the tweets a player receives on the same day before a match on his game performance. 
To preprocess the tweets in order to bring them into a term-document representation (tokenization, stemming, stopword removal, etc.) we will use tidytext.
tidytext provides a set of sentiment lexicons which we will use to assign sentiments to tweets in the following way: 

AFINN from Finn Arup Nielsen [-@IMM2011-06010] provides numerical sentiment scores for a list of words in a range between -5 for strongly negative and +5 for strongly positive statements. We will assign the average AFINN-sentiment score of the terms a respective tweet contains as its sentiment score. 
NRC from Saif Mohammad and Peter Turney [-@noauthor_nrc_nodate] provides categorical sentiments (positive, negative) as well as binary association to emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) to a set of words. We will use the emotions and assign the proportion of terms that correspond to each emotion to a tweet. 

The sentiment scores and emotion proportions of the tweets a player received in the 12-hour window before a game will then be averaged and linked to the respective game statistics of the player which builds our dataset for the analysis. 


<!--**Objective 2: Exploratory Data Analysis**
-->
## Exploratory Data Analysis

For data visualization and to manipulate the data frame with respect to the analysis we will use ggplot2 and dplyr, respectively (both included in tidyverse)
To measure the performance of a player in a game we will use his Plus/Minus score (see above) since it considers the overall contribution of the player to the team when he is on-court (also includes his defensive effort and generally the contribution to his team to score points). 
To assess the relationship between the aggregated sentiment score of tweets a player received on game days and his performance in the games (indicated by the Plus/Minus score) we will perform a regression analysis. 
In case there is no clear linear relationship we may also consider the Predictive Power Score [-@wetschoreck_rip_2020] to assess if there is a non-linear relationship between the two variables. 
We may also distinguish between game days where players received highly positive tweets (e.g. averaged sentiment score above 4) and game days where they received highly negative tweets (e.g. averaged sentiment score below -4) to test if there is a significant difference in game-day performance between these two groups (e.g. by using a two-tailed paired-samples t-test). 

### Sentiment Analysis
The 

<!--**Objective 3: Presentation**

The final project documentation will be written as RMarkdown notebook containing all previously done work and our results. The plots that we created with ggplot2 and dplyr get integrated into the file to underpin our findings.
A project website in which we present our main results as general overview is planned to be written most likely using blogdown combined with an integration of shiny apps, which provides us a good method to embed our results into the web presence.
With OBS Studio as a screen capturing tool and DaVinci Resolve as state-of-the-art video editing software, we plan to record the slideshow that tells a 2-minute summary of the previously mentioned RMarkdown notebook.
The project presentation will then be designed with Microsoft PowerPoint or Prezi which is a powerful tool to create interactive and exciting slideshows.


## Time Plan

**Phase 1: Literature Research about related works, methods and approaches/ Initial project setup** \newline 
Finding relevant literature of related works is significant to identify, whether and to which extent research was already done that is close to our project. On the one hand, this helps to avoid doing research that was already done, on the other hand, we can use the perceptions of this literature as foundation for our work.
Therefore one scope of this phase is to identify related literature and refine our project based on these findings. Furthermore, we then can set up the methods and approaches to use for the coming phases like methods for a well functioning data preparation pipeline, well-suited data analysis approaches or evaluation criteria. \newline 

Steps:

- Finding a research subject that is both, novel and not yet too heavily exploited by other research groups
- Review of related works, methods and approaches that fit our subject and can provide background information, additional insights and useful findings as well as help to sharpen the initial idea of the project
- Retrieval of relevant data that is needed to investigate our research question and checking for suitability (relevant features, data types, how much transformation is needed)
- Formulation of the project proposal \newline 

Milestones/Results:

- Project proposal submission
- Project proposal feedback \newline 

**Phase 2: Data Cleaning, Transformation and Integration** \newline 
Before diving deep into the main task of exploratory data analysis, the relevant data first needs to be extracted, cleaned, transformed and integrated. The main challenge in this phase will be to identify relevant Twitter posts, to prepare them for sentiment analysis and finally extract the attitude of each post. Additionally, the NBA-dataset has to be transformed and integrated to fit with the sentiment polarity scores created beforehand. \newline 

Steps:

- Identification, extraction and cleaning (tokenization, stopword removal, stemming/lemmatization) of relevant Twitter posts via Twitter API
- Sentiment analysis of the extracted posts to generate polarity scores
- Cleaning and transformation of the NBA dataset
- Integration of the preprocessed data sets \newline 

Milestones/Results:

- Final dataset for analysis \newline 

**Phase 3: Exploratory Data Analysis** \newline 
During the main phase of our project, the analysis for possibly existing positive/negative correlations between the extracted Twitter posts and the NBA player performance data is to be done. At first, a descriptive analysis, containing frequencies of variables, min. and max. values, means, etc., shall give insights into the dataset. Then, the data needs to be assessed for relevant variables, expressing the behaviors we want to observe and based on this, the actual analysis can be done.
During the whole process, the results of our research get plotted in expressive figures which then find their way into the final project presentation, website and notebook. \newline 

Steps:

- Descriptive data analysis: Gaining insights into the datasets by frequency analysis, distributions, means, single feature comparison, etc.
- Choice of meaningful predictor and outcome variables that hold the information needed for our correlation analysis
- Identification of possible correlations based on these features
- Creation of meaningful plots to underpin our findings \newline 

Milestones/Results:

- Final correlation analysis \newline 

**Phase 4: Evaluation of the Results** \newline 
To test whether our initial research question holds true, hypothesis testing needs to be done. For this purpose, the correlations found in the phase before need to be checked for significance by using one of several possible approaches. Assessing the proper significance test among Chi-Square test, Regression, T-test, ANOVA test, etc. (to just mention some) is one of the challenges which have to be tackled in this phase of our project. \newline

Steps:

- Testing if and which of the hypotheses that we proclaimed beforehand holds, based on our correlation analysis results 
- Assessing a proper significance test for the given variables
- Statistical testing of the results’ significance \newline

--->
\newpage

## Exploratory Data Analysis

## Final Analysis

## Literature

